Practical Parametricity for GADTs

### 2020-07-13 20-gadts/inf-rules.tex
latest version in progress before laptop repair

To accommodate GADTs, we will need to transition Reynolds' approach
from a $\set$-based semantics to a semantics based on
$\omega$-complete partial orders. We denote the category of
$\omega$-complete partial orders ($\oCPO$s) and their sup-preserving
morphisms by $\omega$CPO. The underlying set of an $\omega$-CPO $A$ is
denoted $|A|$.  The category $\oCPOR$ of relations on $\omega$CPOs has
as its objects triples $(A,B,R)$, where $A, B: \oCPO$ and $R:
\rel(|A|, |B|)$, and has as its morphisms from $(A,B,R)$ to
$(A',B',R')$ pairs $(f : A \to A',g : B \to B')$ of morphisms in
$\oCPO$ such that $(f\,a, g\,b) \in R'$ whenever $(a,b) \in R$. We
write $R : \oCPOR(A,B)$ in place of $(A,B,R): \oCPOR$ when
convenient. If $R : \oCPOR(A,B)$ then we write $\pi_1 R$ and $\pi_2 R$
for the {\em domain} $A$ of $R$ and the {\em codomain} $B$ of $R$,
respectively. If $A : \oCPO$, then we write $\Eq_A = (A,A,\Eq_{|A|})$
for the {\em equality relation} on $A$.

@@

When convenient we identify a functor in $[\oCPO^0, \oCPO]_{\omega_1}$
with its value on $\ast$ and consider a $\oCPO$ environment to map a
type variable of arity $0$ to an $\omega_1$-cocontinuous functor from
$\oCPO^0$ to $\oCPO$, i.e., to an $\oCPO$. If $\Phi =
\{\phi_1^{k_1},...,\phi_n^{k_n}\}$ and $\ol K = \{K_1,...,K_n\}$ where
$K_i : [\oCPO^{k_i},\oCPO]_{\omega_1}$ for $i = 1,...,n$, then we
write $\rho[\ol{\Phi := K}]$ for the $\oCPO$ environment $\rho'$ such
that $\rho' \phi_i = K_i$ for $i = 1,...,n$ and $\rho' \phi = \rho
\phi$ if $\phi \not \in \Phi$.  If $\rho$ is an $\oCPO$ environment,
we write $\Eq_\rho$ for the $\oCPOR$ environment (see
Definition~\ref{def:reln-env}) such that $\Eq_\rho v = \Eq_{\rho v}$
for every type variable $v$. The categories $\oCPORT_k$ and relational
interpretations appearing in the third clause of
Definition~\ref{def:set-sem} are given in full in Section~\ref{?}.

@@

\new{Need to modify to take equality conditions into account.}  If
$\C$ is any category, then given functors $F,G : \C \to \oCPO$ and
natural transformations $\eta, \eta' : F \to G$, we define $\eta \leq
\eta'$ iff $\eta_c \leq \eta'_c$ for all $c \in \C$, i.e., iff
$\eta_c\,x \leq \eta'_c\,x$ in $Gc$ for all $c \in \C$ and $x \in Fc$.
If $(\eta_i)_{i < \omega}$ is a chain in $\{\eta : F \to G\}$, then
the family of morphisms $(\bigvee_{i < \omega} \eta_i)_c = \lambda
x. \bigvee_{i < \omega} (\eta_i)_c\,x : Fc \to Gc$ defines a natural
transformation $\bigvee_{i < \omega} \eta_i : F \to G$, and this
natural transformation is clearly the supremum of $(\eta_i)_{i <
  \omega}$ in $\{\eta : F \to G\}$. We therefore have that $\{\eta : F
\to G\}$ is itself an $\oCPO$.  Letting $F = \lambda \ol
K. \setsem{\Gamma;\Phi \vdash F}\rho[\ol{\Phi := K}]$ and $G = \lambda
\ol K. \setsem{\Gamma;\Phi \vdash G}\rho[\ol{\Phi := K}]$, we have
that $\oCPOsem{\Gamma; \emptyset \vdash \Nat^\Phi F\,G}$ is a
{\color{red} discrete?} functor.
%  indeed an
%  $\oCPO$, i.e., is a constantly-valued functor into $\oCPO$.

Interpretations of $\Nat$-types ensure that $\oCPOsem{\Gamma \vdash F
  \to G}$ and $\oCPOsem{\Gamma \vdash \forall \ol\alpha. F}$ are as
expected in any parametric model.

To make sense of the last clause in Definition~\ref{def:set-sem}, we
need to know that, for each $\rho \in \oCPOenv$, $T^\oCPO_{H,\rho}$ is
an $\omega_1$-cocontinuous endofunctor on $[\oCPO^k,
  \oCPO]_{\omega_1}$, and thus admits a fixpoint.  Since
$T_{H,\rho}^\oCPO$ is defined in terms of
$\oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash H}$, this means
that interpretations of types must be such functors, which in turn
means that the actions of set interpretations of types on objects and
on morphisms in $\oCPOenv$ are intertwined. Fortunately, we know
from~\cite{jp19} that, for every $\Gamma; \Phi \vdash G$,
$\oCPOsem{\Gamma; \Phi \vdash G}$ is actually in
$[\ol{[\oCPO^k,\oCPO]_{\omega_1}},\oCPO]_{\omega_1}$, or,
equivalently, $[\oCPOenv,\oCPO]_{\omega_1}$.  Therefore, for each
$\oCPOsem{\Gamma; \ol\gamma, \phi^k, \ol{\alpha} \vdash H}$, the
corresponding operator $T^\oCPO_{H}$ can be extended to a {\em
  functor} from $\oCPOenv$ to
$[[\oCPO^k,\oCPO]_{\omega_1},[\oCPO^k,\oCPO]_{\omega_1}]_{\omega_1}$.
The action of $T^\oCPO_H$ on an object $\rho \in \oCPOenv$ is given by
the higher-order functor $T_{H,\rho}^\oCPO$, whose actions on objects
(functors in $[\oCPO^k, \oCPO]_{\omega_1}$) and on morphisms (natural
transformations between such functors) are given in
Definition~\ref{def:set-sem}. The action of $T^\oCPO_H$ on a morphism
$f : \rho \to \rho'$ is the higher-order natural transformation
$T^\oCPO_{H,f} : T^\oCPO_{H,\rho} \to T^\oCPO_{H,\rho'}$ whose action
on $F : [\oCPO^k,\oCPO]_{\omega_1}$ is the natural transformation
$T^\oCPO_{H,f}\, F : T^\oCPO_{H,\rho}\,F \to T^\oCPO_{H,\rho'}\,F$
whose component at $\ol A$ is $(T^\oCPO_{H,f}\, F)_{\ol A} =
\oCPOsem{\Gamma; \ol\gamma,\phi,\ol{\alpha} \vdash H}f[\phi :=
  \id_F][\ol{\alpha := \id_A}]$.

In addition, for each $\ol K$, we have that $\textit{Lan}_{\ol K}$ is
itself a (higher-order) functor. Specifically, given functors $F, F' :
\C \to \D$, a sequence of functors $\ol K = K_1, ..., K_n$ with $K_i :
\C \to \C_i$ for $i = 1,...,n$, and a natural transformation $\alpha :
F \to G$, the functorial action $\textit{Lan}_{\ol K} \alpha :
\textit{Lan}_{\ol K} F \to \textit{Lan}_{\ol K} F'$ of
$\mathit{Lan}_{\ol K}$ on $\alpha$ is defined to be the unique natural
transformation such that $((\textit{Lan}_{\ol K} \alpha) \,\circ
\,\langle K_1,...,K_n \rangle) \circ \eta_F = \eta_{F'} \,\circ\,
\alpha$. Here, $\eta_F : F \to (\textit{Lan}_{\ol K} F) \circ \langle
K_1,...,K_n \rangle$ and $\eta_{F'} : F' \to (\textit{Lan}_{\ol K} F')
\circ \langle K_1,...,K_n \rangle$ are the natural transformations
associated with the functors $\textit{Lan}_{\ol K} F$ and
$\textit{Lan}_{\ol K} F'$ from $\Pi_{i \in \{1,...,n\}} \C_i$ to $\D$,
respectively. It is not hard to see that $\mathit{Lan}_{\ol K}$ is a
(higher-order) functor under this definition.

@@

Let $f: \rho \to \rho'$ be a morphism between $\omega$-CPO environments $\rho$

@@

\begin{dfn}\label{def:rel-transf}
A {\em $k$-ary $\oCPO$ relation transformer} $F$ is a triple $(F^1,
F^2,F^*)$, where $F^1,F^2 : [\oCPO^k,\oCPO]_{\omega_1}$ and $F^* :
[\oCPOR^k, \oCPOR]_{\omega_1}$ are such that
$R_1:\oCPOR(A_1,B_1),...,R_k:\oCPOR(A_k,B_k)$ then $F^* \ol{R} :
\oCPOR(F^1 \ol{A}, F^2 \ol{B})$, and if $(\alpha_1, \beta_1) \in
\HomoCPOR(R_1,S_1),..., (\alpha_k, \beta_k) \in \HomoCPOR(R_k,S_k)$
then $F^* \ol{(\alpha, \beta)} = (F^1 \ol{\alpha}, F^2 \ol{\beta})$.
We define $F\ol{R}$ to be $F^*\overline{R}$ and
$F\overline{(\alpha,\beta)}$ to be $F^*\overline{(\alpha,\beta)}$.
\end{dfn}
The last condition of the first sentence of
Definition~\ref{def:rel-transf} expands to: if $\ol{(a,b) \in R}$
implies $\ol{(\alpha\,a,\beta\,b) \in S}$ then $(c,d) \in F^*\ol{R}$
implies $(F^1 \ol{\alpha}\,c,F^2 \ol{\beta}\,d) \in F^*\ol{S}$. When
convenient we identify a $0$-ary $\oCPO$ relation transformer
$(A,B,R)$ with $R : \oCPOR(A,B)$. We may also write $\pi_1 F$ for
$F^1$ and $\pi_2 F$ for $F^2$. We extend these conventions to $\oCPO$
relation environments, introduced in Definition~\ref{def:reln-env}
below, in the obvious way.

@@

Now, for any $k$, any $A : \oCPO$, and any $R : \oCPOR(A, B)$, let
$K^\oCPO_A$ be the constantly $A$-valued functor from $\oCPO^k$ to
$\oCPO$ and $K^\oCPOR_R$ be the constantly $R$-valued functor from
$\oCPOR^k$ to $\oCPOR$.  Also let $0$ denote either the initial object
of $\oCPO$ or the initial object of $\oCPOR$, as appropriate.
Observing that, for every $k$, $K^\oCPO_0$ is initial in
$[\oCPO^k,\oCPO]_{\omega_1}$, and $K^\oCPOR_0$ is initial in
$[\oCPOR^k,\oCPOR]_{\omega_1}$, we have that, for each $k$, $K_0 =
(K^\oCPO_0,K^\oCPO_0,K^\oCPOR_0)$ is initial in $\oCPORT_k$. Thus, if
$T = (T^1,T^2,T^*) : \oCPORT_k \to \oCPORT_k$ is an endofunctor on
$\oCPORT_k$ then we can define the $\omega$-CPO relation transformer
$\mu T$ to be $\colim{n \in \nat}{T^n K_0}$.
It is not hard to see that $\mu T$ is given explicitly as 
\begin{equation}\label{eq:mu}
  \mu T = (\mu T^1,\mu T^2,
\colim{n \in \nat}{(T^nK_0)^*})
\end{equation}
and that, as our notation suggests, it really is a fixpoint for $T$ if
$T$ is $\omega_1$-cocontinuous:
\begin{lemma}\label{lem:fp}
For any $T : [\oCPORT_k,\oCPORT_k]_{\omega_1}$, $\mu T \cong T(\mu T)$.
\end{lemma}
\noindent
The isomorphism is given by the morphisms $(\mathit{in}_1,
\mathit{in}_2) : T(\mu T) \to \mu T$ and $(in_1^{-1}, in_2^{-1}) : \mu
T \to T(\mu T)$ in $\oCPORT_k$. The latter is always a morphism in $\oCPORT_k$,
but the former need not be if $T$ is not $\omega_1$-cocontinuous.

@@

{\color{red} HERE!!!!!}

@@

\begin{dfn}\label{def:rel-sem}
The {\em relational interpretation}
$\oCPORsem{\cdot} : \F \to [\oCPORenv, \oCPOR]_{\omega_1}$ is defined by
\begin{align*}
  \oCPORsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \oCPORsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
  \oCPORsem{\Gamma; \emptyset \vdash \Nat^{\ol\psi} \,F\,G}\rho &= \{\eta
  : \lambda \ol{K}.\,\oCPORsem{\Gamma; \ol{\psi} \vdash
    F}\rho[\ol{\psi := K}] \Rightarrow \lambda \ol{K}. \,\oCPORsem{
    \Gamma; \ol{\psi} \vdash G}\rho[\ol{\psi := K}]\}\\
  &=
  \{(t,t') \in \oCPOsem{\Gamma; \emptyset
    \vdash \Nat^{\ol{\psi}}
    \,F\,G} (\pi_1 \rho) \times \oCPOsem{ 
    \Gamma;\emptyset
    \vdash \Nat^{\ol{\psi}} \,F\,G} (\pi_2
  \rho)~|~\\ 
  & \hspace{0.3in} \forall \ol{K = (K^0, K^1, K^*) : \oCPORT_{k}}.\\
  & \hspace{0.4in} (t_{\ol{K^0}},t'_{\ol{K^1}}) \in
  (\oCPORsem{\Gamma; \ol{\psi} \vdash G}\rho[\ol{\psi :=
      K}])^{\oCPORsem{\Gamma;\ol{\psi}\vdash F}\rho[\ol{\psi := K}]} \}\\  
  \oCPORsem{\Gamma;\Phi \vdash \phi \ol{F}}\rho &=
  (\rho\phi)\ol{\oCPORsem{\Gamma;\Phi \vdash 
    F}\rho}\\
  \oCPORsem{\Gamma;\Phi \vdash F+G}\rho &=
  \oCPORsem{\Gamma;\Phi \vdash F}\rho +
  \oCPORsem{\Gamma;\Phi \vdash G}\rho\\
  \oCPORsem{\Gamma;\Phi \vdash F\times G}\rho &=
  \oCPORsem{\Gamma;\Phi \vdash F}\rho \times
  \oCPORsem{\Gamma;\Phi \vdash G}\rho\\  
   \oCPORsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho
  &= (\mu T_{H,\rho})\ol{\oCPORsem{\Gamma;\Phi,\ol\gamma \vdash
     G}\rho}\\
  \text{where }	T_{H,\rho}
    &= (T^\set_{H,\pi_1\rho}, T^\set_{H,\pi_2\rho}, T^\rel_{H,\rho}) \\
  \text{and } T^\oCPOR_{H,\rho}\,F
    &= \lambda \ol{R}. \oCPORsem{
      \Gamma;\ol\gamma,\phi,\ol{\alpha} \vdash H}\rho[\phi :=
    F][\ol{\alpha := R}]\\
  \text{and } T^\oCPOR_{H,\rho}\,\delta
    &= \lambda \ol{R}. \oCPORsem{
      \Gamma;\ol\gamma,\phi,\ol{\alpha} \vdash H}\id_\rho[\phi :=
    \delta][\ol{\alpha := \id_{\ol{R}}}] \\
  \oCPORsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\rho &=
    (\textit{Lan}_{\ol{\oCPORsem{\Gamma; \ol\alpha \vdash K}\rho[\ol{\alpha := \_}]}}
      \oCPORsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho[\ol{\alpha := \_}])
    \ol{\oCPORsem{\Gamma; \Phi \vdash A}\rho}
\end{align*}
\end{dfn}

>>>>

To accommodate GADTs, we will need to transition Reynolds' approach
from a $\set$-based semantics to a semantics based on
$\omega$-complete partial orders. We denote the category of
$\omega$-complete partial orders ($\oCPO$s) and their sup-preserving
morphisms by $\omega$CPO. The underlying set of an $\oCPO$ $A$ is
denoted $|A|$.  The category $\oCPOR$ of relations on $\omega$CPOs has
as its objects triples $(A,B,R)$, where $A, B: \oCPO$ and $R:
\rel(|A|, |B|)$, and has as its morphisms from $(A,B,R)$ to
$(A',B',R')$ pairs $(f : A \to A',g : B \to B')$ of morphisms in
$\oCPO$ such that $(f\,a, g\,b) \in R'$ whenever $(a,b) \in R$. We
write $R : \oCPOR(A,B)$ in place of $(A,B,R): \oCPOR$ when
convenient. If $R : \oCPOR(A,B)$ then we write $\pi_1 R$ and $\pi_2 R$
for the {\em domain} $A$ of $R$ and the {\em codomain} $B$ of $R$,
respectively. If $A : \oCPO$, then we write $\Eq_A = (A,A,\Eq_{|A|})$
for the {\em equality relation} on $A$.

@@

When convenient we identify a functor in $[\oCPO^0, \oCPO]_{\omega_1}$
with its value on $\ast$ and consider a $\oCPO$ environment to map a
type variable of arity $0$ to an $\omega_1$-cocontinuous functor from
$\oCPO^0$ to $\oCPO$, i.e., to an $\oCPO$. If $\Phi =
\{\phi_1^{k_1},...,\phi_n^{k_n}\}$ and $\ol K = \{K_1,...,K_n\}$ where
$K_i : [\oCPO^{k_i},\oCPO]_{\omega_1}$ for $i = 1,...,n$, then we
write $\rho[\ol{\Phi := K}]$ for the $\oCPO$ environment $\rho'$ such
that $\rho' \phi_i = K_i$ for $i = 1,...,n$ and $\rho' \phi = \rho
\phi$ if $\phi \not \in \Phi$.  If $\rho$ is an $\oCPO$ environment,
we write $\Eq_\rho$ for the $\oCPOR$ environment (see
Definition~\ref{def:reln-env}) such that $\Eq_\rho v = \Eq_{\rho v}$
for every type variable $v$. The categories $\oCPORT_k$ and relational
interpretations appearing in the third clause of
Definition~\ref{def:set-sem} are given in full in
Section~\ref{sec:iel}.

{\color{red} Need to describe $\omega_1$-directed colimits in
  $\oCPOenv$ for the type of $\oCPOsem{\cdot}$ to make sense. We do it
  pointwise.}

@@

\new{Need to modify to take equality conditions into account.}  If
$\C$ is any category, then given functors $F,G : \C \to \oCPO$ and
natural transformations $\eta, \eta' : F \to G$, we define $\eta \leq
\eta'$ iff $\eta_c \leq \eta'_c$ for all $c \in \C$, i.e., iff
$\eta_c\,x \leq \eta'_c\,x$ in $Gc$ for all $c \in \C$ and $x \in Fc$.
If $(\eta_i)_{i < \omega}$ is a chain in $\{\eta : F \to G\}$, then
the family of morphisms $(\bigvee_{i < \omega} \eta_i)_c = \lambda
x. \bigvee_{i < \omega} (\eta_i)_c\,x : Fc \to Gc$ defines a natural
transformation $\bigvee_{i < \omega} \eta_i : F \to G$, and this
natural transformation is clearly the supremum of $(\eta_i)_{i <
  \omega}$ in $\{\eta : F \to G\}$. We therefore have that $\{\eta : F
\to G\}$ is itself an $\oCPO$.  Letting $F = \lambda \ol
K. \setsem{\Gamma;\Phi \vdash F}\rho[\ol{\Phi := K}]$ and $G = \lambda
\ol K. \setsem{\Gamma;\Phi \vdash G}\rho[\ol{\Phi := K}]$, we have
that $\oCPOsem{\Gamma; \emptyset \vdash \Nat^\Phi F\,G}\rho$ is indeed
an $\oCPO$. \new{Does this use functorial actions of interps? I don't
  think so, but double-check with Enrico.} {\color{red} That each of
  the above interpretations is $\omega_1$-cocontinuous follows from
  Corollary~12 of~\cite{jp19} if we APPROPRIATELY RESTRICT THE
  SUBSCRIPTS OF $\Lan$s.  For $\Nat$-types, we note that
  $\oCPOsem{\Gamma; \emptyset \vdash \Nat^\Phi F\,G}$ is an
  $\omega_1$-cocontinuous functor because it is constant on
  $\omega_1$-directed sets.} Interpretations of $\Nat$-types ensure
that $\oCPOsem{\Gamma \vdash F \to G}$ and $\oCPOsem{\Gamma \vdash
  \forall \ol\alpha. F}$ are as expected in any parametric model.

To make sense of the next-to-last clause in
Definition~\ref{def:set-sem}, we need to know that, for each $\rho \in
\oCPOenv$, $T^\oCPO_{H,\rho}$ is an $\omega_1$-cocontinuous
endofunctor on $[\oCPO^k, \oCPO]_{\omega_1}$, and thus admits a
fixpoint.  Since $T_{H,\rho}^\oCPO$ is defined in terms of
$\oCPOsem{\Gamma;\ol\gamma,\phi, \ol{\alpha} \vdash H}$, this means
that interpretations of types must be such functors, which in turn
means that the actions of set interpretations of types on objects and
on morphisms in $\oCPOenv$ are intertwined. Fortunately, we know
from~\cite{jp19} that, for every $\Gamma; \ol\alpha \vdash F$,
$\oCPOsem{\Gamma; \Phi \vdash F}$ is actually in
$[\oCPO^k,\oCPO]_{\omega_1}$, where $k = |\ol\alpha|$.  Therefore, for
each $\oCPOsem{\Gamma; \ol\gamma, \phi^k, \ol{\alpha} \vdash H}$, the
corresponding operator $T^\oCPO_{H}$ can be extended to a {\em
  functor} from $\oCPOenv$ to
$[[\oCPO^k,\oCPO]_{\omega_1},[\oCPO^k,\oCPO]_{\omega_1}]_{\omega_1}$.
The action of $T^\oCPO_H$ on an object $\rho \in \oCPOenv$ is given by
the higher-order functor $T_{H,\rho}^\oCPO$, whose actions on objects
(functors in $[\oCPO^k, \oCPO]_{\omega_1}$) and on morphisms (natural
transformations between such functors) are given in
Definition~\ref{def:set-sem}. The action of $T^\oCPO_H$ on a morphism
$f : \rho \to \rho'$ is the higher-order natural transformation
$T^\oCPO_{H,f} : T^\oCPO_{H,\rho} \to T^\oCPO_{H,\rho'}$ whose action
on $F : [\oCPO^k,\oCPO]_{\omega_1}$ is the natural transformation
$T^\oCPO_{H,f}\, F : T^\oCPO_{H,\rho}\,F \to T^\oCPO_{H,\rho'}\,F$
whose component at $\ol A$ is $(T^\oCPO_{H,f}\, F)_{\ol A} =
\oCPOsem{\Gamma; \ol\gamma,\phi,\ol{\alpha} \vdash H}f[\phi :=
  \id_F][\ol{\alpha := \id_A}]$.

%For the last clause in Definition~\ref{def:set-sem} to be well-defined
%we need $\oCPOsem{\Gamma; \ol\alpha \vdash K}\rho[\ol{\alpha := \_}]$
%and $\oCPOsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho[\ol{\alpha :=
%    \_}]$ to be $\omega_1$-cocontinuous functors.
In addition, for each $\ol K$, we have that $\textit{Lan}_{\ol K}$ is
itself a (higher-order) functor. Specifically, given functors $F, F' :
\C \to \D$, a sequence of functors $\ol K = K_1, ..., K_n$ with $K_i :
\C \to \C_i$ for $i = 1,...,n$, and a natural transformation $\alpha :
F \to G$, the functorial action $\textit{Lan}_{\ol K} \alpha :
\textit{Lan}_{\ol K} F \to \textit{Lan}_{\ol K} F'$ of
$\mathit{Lan}_{\ol K}$ on $\alpha$ is defined to be the unique natural
transformation such that $((\textit{Lan}_{\ol K} \alpha) \,\circ
\,\langle K_1,...,K_n \rangle) \circ \eta_F = \eta_{F'} \,\circ\,
\alpha$. Here, $\eta_F : F \to (\textit{Lan}_{\ol K} F) \circ \langle
K_1,...,K_n \rangle$ and $\eta_{F'} : F' \to (\textit{Lan}_{\ol K} F')
\circ \langle K_1,...,K_n \rangle$ are the natural transformations
associated with the functors $\textit{Lan}_{\ol K} F$ and
$\textit{Lan}_{\ol K} F'$ from $\Pi_{i \in \{1,...,n\}} \C_i$ to $\D$,
respectively. It is not hard to see that $\mathit{Lan}_{\ol K}$ is a
(higher-order) functor under this definition.

@@

Let $f: \rho \to \rho'$ be a morphism between $\oCPO$ environments $\rho$

@@

\begin{dfn}\label{def:rel-transf}
A {\em $k$-ary $\oCPO$ relation transformer} $F$ is a triple $(F^1,
F^2,F^*)$, where $F^1,F^2 : [\oCPO^k,\oCPO]_{\omega_1}$ and $F^* :
[\oCPOR^k, \oCPOR]_{\omega_1}$ are such that if
$R_1:\oCPOR(A_1,B_1),...,R_k:\oCPOR(A_k,B_k)$ then $F^* \ol{R} :
\oCPOR(F^1 \ol{A}, F^2 \ol{B})$, and if $(\alpha_1, \beta_1) \in
\HomoCPOR(R_1,S_1),..., (\alpha_k, \beta_k) \in \HomoCPOR(R_k,S_k)$
then $F^* \ol{(\alpha, \beta)} = (F^1 \ol{\alpha}, F^2 \ol{\beta})$.
We define $F\ol{R}$ to be $F^*\overline{R}$ and
$F\overline{(\alpha,\beta)}$ to be $F^*\overline{(\alpha,\beta)}$.
\end{dfn}
The last condition of the first sentence of
Definition~\ref{def:rel-transf} expands to: if $\ol{(a,b) \in R}$
implies $\ol{(\alpha\,a,\beta\,b) \in S}$ then $(c,d) \in F^*\ol{R}$
implies $(F^1 \ol{\alpha}\,c,F^2 \ol{\beta}\,d) \in F^*\ol{S}$. When
convenient we identify a $0$-ary $\oCPO$ relation transformer
$(A,B,R)$ with $R : \oCPOR(A,B)$. We may also write $\pi_1 F$ for
$F^1$ and $\pi_2 F$ for $F^2$. We extend these conventions to $\oCPO$
relation environments, introduced in Definition~\ref{def:reln-env}
below, in the obvious way.

@@

Now, for any $k$, any $A : \oCPO$, and any $R : \oCPOR(A, B)$, let
$K^\oCPO_A$ be the constantly $A$-valued functor from $\oCPO^k$ to
$\oCPO$ and $K^\oCPOR_R$ be the constantly $R$-valued functor from
$\oCPOR^k$ to $\oCPOR$.  Also let $0$ denote either the initial object
of $\oCPO$ or the initial object of $\oCPOR$, as appropriate.
Observing that, for every $k$, $K^\oCPO_0$ is initial in
$[\oCPO^k,\oCPO]_{\omega_1}$, and $K^\oCPOR_0$ is initial in
$[\oCPOR^k,\oCPOR]_{\omega_1}$, we have that, for each $k$, $K_0 =
(K^\oCPO_0,K^\oCPO_0,K^\oCPOR_0)$ is initial in $\oCPORT_k$. Thus, if
$T = (T^1,T^2,T^*) : \oCPORT_k \to \oCPORT_k$ is an endofunctor on
$\oCPORT_k$ then we can define the $\oCPO$ relation transformer
$\mu T$ to be $\colim{n \in \nat}{T^n K_0}$.
It is not hard to see that $\mu T$ is given explicitly as 
\begin{equation}\label{eq:mu}
  \mu T = (\mu T^1,\mu T^2,
\colim{n \in \nat}{(T^nK_0)^*})
\end{equation}
and that, as our notation suggests, it really is a fixpoint for $T$ if
$T$ is $\omega_1$-cocontinuous:
\begin{lemma}\label{lem:fp}
For any $T : [\oCPORT_k,\oCPORT_k]_{\omega_1}$, $\mu T \cong T(\mu T)$.
\end{lemma}
\noindent
The isomorphism is given by the morphisms $(\mathit{in}_1,
\mathit{in}_2) : T(\mu T) \to \mu T$ and $(in_1^{-1}, in_2^{-1}) : \mu
T \to T(\mu T)$ in $\oCPORT_k$. The latter is always a morphism in $\oCPORT_k$,
but the former need not be if $T$ is not $\omega_1$-cocontinuous.

@@

@@

\begin{dfn}\label{def:rel-sem}
The {\em relational interpretation}
$\oCPORsem{\cdot} : \F \to [\oCPORenv, \oCPOR]_{\omega_1}$ is defined by
  \oCPORsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \oCPORsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
  \oCPORsem{\Gamma; \emptyset \vdash \Nat^\Phi \,F\,G}\rho &= \{\eta
  : \lambda \ol{K}.\,\oCPORsem{\Gamma; \Phi \vdash
    F}\rho[\ol{\Phi := K}] \Rightarrow \lambda \ol{K}. \,\oCPORsem{
    \Gamma; \Phi \vdash G}\rho[\ol{\phi := K}]\}\\
  &=
  \{(t,t') \in \oCPOsem{\Gamma; \emptyset
    \vdash \Nat^\Phi
    \,F\,G} (\pi_1 \rho) \times \oCPOsem{ 
    \Gamma;\emptyset
    \vdash \Nat^\Phi \,F\,G} (\pi_2
  \rho)~|~\\ 
  & \hspace{0.3in} \forall \ol{K = (K^0, K^1, K^*) : \oCPORT_{k}}.\\
  & \hspace{0.4in} (t_{\ol{K^0}},t'_{\ol{K^1}}) \in
  (\oCPORsem{\Gamma; \Phi \vdash G}\rho[\ol{\Phi :=
      K}])^{\oCPORsem{\Gamma;\Phi\vdash F}\rho[\ol{\Phi := K}]} \}\\  
  \oCPORsem{\Gamma;\Phi \vdash \phi \ol{F}}\rho &=
  (\rho\phi)\ol{\oCPORsem{\Gamma;\Phi \vdash 
    F}\rho}\\
  \oCPORsem{\Gamma;\Phi \vdash F+G}\rho &=
  \oCPORsem{\Gamma;\Phi \vdash F}\rho +
  \oCPORsem{\Gamma;\Phi \vdash G}\rho\\
  \oCPORsem{\Gamma;\Phi \vdash F\times G}\rho &=
  \oCPORsem{\Gamma;\Phi \vdash F}\rho \times
  \oCPORsem{\Gamma;\Phi \vdash G}\rho\\  
   \oCPORsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho
  &= (\mu T_{H,\rho})\ol{\oCPORsem{\Gamma;\Phi,\ol\gamma \vdash
     G}\rho}\\
  \text{where }	T_{H,\rho}
    &= (T^\set_{H,\pi_1\rho}, T^\set_{H,\pi_2\rho}, T^\rel_{H,\rho}) \\
  \text{and } T^\oCPOR_{H,\rho}\,F
    &= \lambda \ol{R}. \oCPORsem{
      \Gamma;\ol\gamma,\phi,\ol{\alpha} \vdash H}\rho[\phi :=
    F][\ol{\alpha := R}]\\
  \text{and } T^\oCPOR_{H,\rho}\,\delta
    &= \lambda \ol{R}. \oCPORsem{
      \Gamma;\ol\gamma,\phi,\ol{\alpha} \vdash H}\id_\rho[\phi :=
    \delta][\ol{\alpha := \id_{\ol{R}}}] \\
  \oCPORsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\rho &=
    (\textit{Lan}_{\ol{\oCPORsem{\Gamma; \ol\alpha \vdash K}\rho[\ol{\alpha := \_}]}}
      \oCPORsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho[\ol{\alpha := \_}])
    \ol{\oCPORsem{\Gamma; \Phi \vdash A}\rho}
The interpretations in Definitions~\ref{def:rel-sem}
and~\ref{def:rel-sem-funcs} below respect weakening, and also ensure
that $\oCPORsem{\Gamma \vdash F \to G}$ and $\oCPORsem{\Gamma \vdash
  \forall \ol \alpha. F}$ are as expected in any parametric model.
Observe that $\oCPORsem{\Gamma; \emptyset \vdash \Nat^\Phi F\,G} \rho$
is in $\oCPOR$. \new{Same comments here as above.} {\color{red} It is
  indeed a relation between $\oCPO$s, and it preserves sups of
  $\omega$-chains of natural transformations because their sups are
  computed pointwise. Concretely, if $(t_i, t_i') \colon
  \oCPORsem{\Gamma; \emptyset \vdash \Nat^\Phi F\,G} \rho$ for all $i
  < \omega$, then for all $\ol{K = (K^0, K^1, K^*) : \oCPORT_{k}}$ we
  have that
\[((t_i)_{\ol{K^0}},\, (t'_i)_{\ol{K^1}}) \in (\oCPORsem{\Gamma; \Phi
  \vdash G}\rho[\ol{\Phi := K}])^{\oCPORsem{\Gamma;\Phi\vdash
    F}\rho[\ol{\Phi := K}]}\] for every $i < \omega$. In addition,
$((t_i)_{\ol{K^0}})_{i < \omega}$ is an $\omega$-chain in
$(\oCPOsem{\Gamma; \Phi \vdash G}\rho[\ol{\Phi :=
    K^0}])^{\oCPOsem{\Gamma;\Phi\vdash F}\rho[\ol{\Phi := K^0}]}$ and
$((t'_i)_{\ol{K^1}})_{i < \omega}$ is an $\omega$-chain in
$(\oCPOsem{\Gamma; \Phi \vdash G}\rho[\ol{\Phi :=
    K^1}])^{\oCPOsem{\Gamma;\Phi\vdash F}\rho[\ol{\Phi := K^1}]}$, so
their sups are not only well-defined but must also be related
in
$(\oCPORsem{\Gamma;
    \Phi \vdash G}\rho[\ol{\Phi := K}])^{\oCPORsem{\Gamma;\Phi\vdash
      F}\rho[\ol{\Phi := K}]}$. This gives that
  \Big((\bigvee_{i < \omega} t_i)_{\ol{K^0}}, (\bigvee_{i < \omega}
  t'_i)_{\ol{K^1}} \Big) = \Big(\bigvee_{i < \omega} (t_i)_{\ol{K^0}},
  \bigvee_{i < \omega} (t'_i)_{\ol{K^1}} \Big) \in (\oCPORsem{\Gamma;
    \Phi \vdash G}\rho[\ol{\Phi := K}])^{\oCPORsem{\Gamma;\Phi\vdash
      F}\rho[\ol{\Phi := K}]}
Moreover, as for $\oCPO$ interpretations, $\omega_1$-cocontinuity of
each of the above interpretations follows from Corollary~12
of~\cite{jp19} if we APPROPRIATELY RESTRICT THE SUBSCRIPTS of the
$\Lan$s. For $\Nat$-types, $\oCPORsem{\Gamma; \emptyset \vdash
  \Nat^\Phi F\,G}$ is an $\omega_1$-cocontinuous functor because it is
constant on $\omega_1$-directed sets.} If $\rho \in \oCPORenv$ and
$\vdash F$, then we write $\oCPORsem{\vdash F}$ instead of
$\oCPORsem{\vdash F}\rho$.  For the second-to-last clause in
Definition~\ref{def:rel-sem} to be well-defined we need $T_{H,\rho}$
to be an $\omega_1$-cocontinuous endofunctor on $\oCPORT$ so that, by
